Overrelaxed Sinkhorn--Knopp Algorithm for Regularized Optimal Transport

### 2017-10-20 paper.tex
Relecture Nicolas

Crucial properties of this Lyapunov function are gathered in the next lemma.
\begin{lemma} \label{lemma:lyapunov_decrease}
	For any $M \in \IR_+^*$, the sublevel set $\left\{ \gamma \mid F(\gamma) \le M \right\}$ is compact.
%\end{lemma}
%\begin{lemma}
	Moreover, for $\gamma$ in $\IR^{mn}_{+*}$, the decrease of the Lyapunov function after an overrelaxed projection can be computed as
	\begin{equation} \label{eq:kl_diff_scal}
	F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma)) = 
	\scal{\mu^k}{\varphi_\omega \left((A_k \gamma) \oslash \mu^k \right)},
	\end{equation}
	where
	\begin{equation}
	\varphi_\omega(x) = x(1-x^{-\omega}) - \omega \log x
	\end{equation}
	is a real function, applied coordinate-wise.
\end{lemma}

\begin{proof}
The fact that the Kullback-Leibler divergence is jointly lower semicontinuous implies in particular that $K$ is closed. Moreover, $K\subset \mathbb{R}^{n_1\times n_2}_+$ is bounded because $F$ is the sum of nonnegative, coercive functions of each component of its argument $\gamma$.
%\end{proof}
%\begin{proof}
%
Formula \eqref{eq:kl_diff_scal} comes the expression $F(\gamma^1)-F(\gamma^2)= \sum_{i,j}\left(\gamma^*_{i,j}\log(\gamma^2_{i,j}/\gamma^1_{i,j})+\gamma^1_{i,j}-\gamma^2_{i,j}\right)$ and the relations \eqref{eq:def_or_proj} and \eqref{scaling}.
\end{proof}

@@

It follows from Lemma \ref{lemma:lyapunov_decrease} that the decrease of $F$ for an overrelaxed projection is very cheap to estimate, since its computational cost is linear with respect to the dimension of data $\mu^k$. In Figure \ref{phi_omega}, we display the function  $\varphi_\omega(x)$. Notice that for the SK algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always nonnegative. For other values $1\le\omega<2$, it is nonnegative for $x$ close to 1.

@@

The resulting algorithm, which is proved to be convergent by Theorem~\ref{thm:algo}, is written in pseudo-code in Algorithm \ref{SOR}. It uses the function $\Theta$ defined implicitly in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which decreasing is as can be seen on Figure~\ref{phi_omega}. Note that with the choice $\theta_0=1$, one recovers exactly the original SK algorithm.

>>>>

Crucial properties of this Lyapunov function are gathered in the next lemma.
\begin{lemma} \label{lemma:lyapunov_decrease}
	For any $M \in \IR_+^*$, the sublevel set $\left\{ \gamma \mid F(\gamma) \le M \right\}$ is compact.
%\end{lemma}
%\begin{lemma}
	Moreover, for any $\gamma$ in $\IR^{mn}_{+*}$, the decrease of the Lyapunov function after an overrelaxed projection can be computed as
	\begin{equation} \label{eq:kl_diff_scal}
	F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma)) = 
	\scal{\mu^k}{\varphi_\omega \left((A_k \gamma) \oslash \mu^k \right)},
	\end{equation}
	where
	\begin{equation}
	\varphi_\omega(x) = x(1-x^{-\omega}) - \omega \log x
	\end{equation}
	is a real function, applied coordinate-wise.
\end{lemma}

\begin{proof}
The fact that the Kullback-Leibler divergence is jointly lower semicontinuous implies in particular that $K$ is closed. Moreover, $K\subset \mathbb{R}^{n_1\times n_2}_+$ is bounded because $F$ is the sum of nonnegative, coercive functions of each component of its argument $\gamma$.
%\end{proof}
%\begin{proof}

Formula \eqref{eq:kl_diff_scal} comes the expression $F(\gamma^1)-F(\gamma^2)= \sum_{i,j}\left(\gamma^*_{i,j}\log(\gamma^2_{i,j}/\gamma^1_{i,j})+\gamma^1_{i,j}-\gamma^2_{i,j}\right)$ and the relations \eqref{eq:def_or_proj} and \eqref{scaling}.
\end{proof}

@@

It follows from Lemma \ref{lemma:lyapunov_decrease} that the decrease of $F$ for an overrelaxed projection is very cheap to estimate, since its computational cost is linear with respect to the dimension of data $\mu^k$. In Figure \ref{phi_omega}, we display the function  $\varphi_\omega(x)$. Notice that for the Sinkhorn--Knopp algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always nonnegative. For other values $1\le\omega<2$, it is nonnegative for $x$ close to 1.

@@

The resulting algorithm, which is proved to be convergent by Theorem~\ref{thm:algo}, is written in pseudo-code in Algorithm \ref{SOR}. It uses the function $\Theta$ defined implicitly in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which decreasing is as can be seen on Figure~\ref{phi_omega}. With the choice $\theta_0=1$, one recovers exactly the original SK algorithm.
